---
title: Accounting for detection bias increases trait-matching in a tropical plant-hummingbird network
author: "Ben Weinstein - Stony Brook University"
date: "Monday, March 16, 2015"
output: html_document
---

```{r,warning=FALSE,message=FALSE,echo=FALSE,cache=FALSE}
library(reshape2)
library(chron)
library(ggplot2)
library(knitr)
library(R2jags)
library(dplyr)
library(stringr)
library(gridExtra)
library(boot)
library(picante)
library(bipartite)

opts_chunk$set(message=FALSE,warning=FALSE,fig.width=5,fig.height=4,echo=TRUE,cache=TRUE,cache.path = 'jp_cache/',fig.align='center',fig.path="figure/")

setwd("C:/Users/Ben/Documents/Occupy/")

set.seed(3)
#source functions

source("Functions.R")
#load("Abundance.RData")
```

#Simulation   

I first create data with constant detection for each hummingbird speices, and the interaction frequency is a function of a species hierarcichal model and corolla similarity. There is uneven sampling among species.

### Parameters

* 10 hummingbird species
* Range of hummingbird bill sizes (mm) ~ Pois(2)
* Twenty plants
* Range of corolla sizes (mm) ~ Pois(2)
* Mean frequeny ($\lambda$) for each hummingbird is drawn from U(0,10)  
* For each plant the occupancy is N($\lambda$,0.2) (truncated 0-1) 
* Trait matching (minimizing Bill-Corolla difference) is drawn from a hierarcichal distribution
$$log(\lambda)<-\alpha_i + \beta_i *traitmatch$$
$$\alpha=N(3,.01)$$
$$\beta = N(-.01,.001)$$

* Imperfect detection 
* $p_i = U(0,1)$ 
* 24 month replicates
* Phenology = .5 (plants are in flower/present .5 of surveys)

**View simulated strength and form of trait matching **

```{r,fig.height=5,fig.width=8}
#Number of hummingbird species
h_species=10
plant_species=20
months=24
detection=runif(h_species,0,1)
phenology=.75

#Bill sizes
Bill<-rpois(h_species,10)

#Corolla sizes
Corolla<-rpois(plant_species,15)

#Subtract both and take absolute value
traitmatch<-abs(sapply(Corolla,function(x) x - Bill)/10)

#regression slope
gamma=-0.5
intercept<-3
sigma_slope<- 0.15
sigma_intercept<- 0.15

beta<-rnorm(h_species,gamma,sigma_slope)
alpha<-rnorm(h_species,intercept,sigma_intercept)

#fit regression
lambda<-exp(alpha + beta * traitmatch)

true_interactions<-sapply(lambda,function(x){rpois(1,x)})

#convert to matrix
true_interactions<-matrix(nrow=h_species,ncol=plant_species,data=true_interactions)

#combine and melt into a single dataframe
mdat<-dcast(melt(list(y=true_interactions,x=traitmatch)),Var1+Var2~L1)

ggplot(mdat,aes(x=x,y=y,col=as.factor(Var1))) + geom_point() + geom_smooth(aes(group=1),method="glm",family="poisson",formula=y~x) + theme_bw() + labs(x="Bill-Corolla Difference",y="Occurrences","Hummingbird") + ggtitle("Correlation in Simulated Data") + labs(col="Species")

```

### True Interaction Matrix

```{r,fig.width=12}
#Reshape into a nicer format
colnames(mdat)<-c("Hummingbird","Plant","Bill_Diff","True_State")

trueplot<-ggplot(mdat,aes(y=as.factor(Plant),x=as.factor(Hummingbird),fill=True_State)) + geom_tile() + labs(x="Hummingbird",y="Plant",fill="Presence") + ggtitle("True State ") + scale_fill_continuous(low="white",high="red","Occurrences")
```

```{r,fig.width=11,fig.height=4}
##Simulate detection 

#for each species loop through and create a replicate dataframe
obs<-array(dim=c(h_species,plant_species,months))

#Which months are plants in bloom?
#This is the phenology matrix
sampled<-replicate(plant_species,rbinom(months,1,phenology))
sampled[sampled==0]<-NA

for(x in 1:h_species){
  for (y in 1:plant_species){
    
    #True State
    ts<-true_interactions[x,y]
    
    #detections are function of phenology detection probability and occupancy
    det<-rbinom(months,ts,detection[x])
    
    #detections and phenology
    obs[x,y,]<-det *sampled[,y]
    }
  }

#plot observed state
obs.state<-melt(obs)
colnames(obs.state)<-c("Hummingbird","Plant","Month","Detections")

#turn NA's to 0
obs.state$Detections[is.na(obs.state$Detections)]<-0

obs.state$Hummingbird<-as.factor(obs.state$Hummingbird)
obs.state$Plant<-factor(obs.state$Plant,levels=1:plant_species)

ob1<-ggplot(obs.state[obs.state$Month==12,],aes(x=Hummingbird,y=Plant,fill=Detections)) + geom_tile() + scale_fill_continuous(low="white",high="Red",limits=c(0,max(mdat$True_State))) + ylab("Plant") + ggtitle("Example 1") + xlab("Hummingbird")

ob2<-ggplot(obs.state[obs.state$Month==18,],aes(x=Hummingbird,y=Plant,fill=Detections)) + geom_tile() + scale_fill_continuous(low="white",high="Red",limits=c(0,max(mdat$True_State))) + ylab("Plant") + ggtitle("Example 2") + xlab("Hummingbird")

grid.arrange(trueplot,ob1,ob2,nrow=1)
```

The left hand panel is the true state for each hummingbird plant combination. The middle and right hand panel are two realizations that might result from a month of sampling given incomplete detection and phenology.

```{r}
d<-merge(mdat,obs.state)
ggplot(d,aes(x=True_State,y=Detections,col=Month)) + geom_point() + theme_bw() + geom_abline() + coord_equal()
```

### Model Fitting
## Hierarcichal Occupancy Model Formulation

$$ Y_{i,j,k} \sim Binom(N_{i,j},detect_i)$$
$$N_{i,j} \sim Pois(\lambda_{i,j}) $$
$$log(\lambda_{i,j})<-\alpha_i + \beta_i * abs(Bill_i - Corolla_i) $$
$$detect_i \sim U(0,1)$$     

**Priors**

$$\alpha_i \sim N(intercept,\tau_{\alpha})$$
$$\beta_i \sim N(\gamma,\tau_{\beta})$$

**Hyperpriors**
$$gamma \sim N(0.001,0.001)$$
$$intercept \sim N(0.001,0.001)$$

$$\tau_{\alpha} \sim Gamma(0.001,0.001)$$
$$\tau_\beta \sim Gamma(0.001,0.001)$$

**Derived quantities**

$$\sigma_{int} = \frac{1}{\tau_{\alpha}}^2$$
$$\sigma_{slope} = \frac{1}{\tau_{\beta}}^2$$

##Hierarcichal Bayesian Model without detection

Detect parameters is 'fixed' by placing it centered on 1 with extremely low variance.

```{r,eval=T}
setwd("C:/Users/Ben/Documents/Occupy/")
runs<-20000

#Source model
source("Bayesian/NoDetect.R")

#print model
print.noquote(readLines("Bayesian/NoDetect.R"))

#Input Data
Dat <- list(
  Y=obs,
  Birds=dim(obs)[1],
  Plants=dim(obs)[2],
  Months=dim(obs)[3],
  traitmatch=traitmatch)

#A blank Y matrix - all present
initY<-array(dim=c(Dat$Birds,Dat$Plants),data=max(obs,na.rm=T))

#Inits
InitStage <- function() {list(beta=rep(.5,Dat$Birds),alpha=rep(.5,Dat$Birds),detect=rep(.999,Dat$Birds),intercept=0,tau_alpha=0.1,tau_beta=0.1,N=initY,gamma=0)}

#Parameters to track
ParsStage <- c("detect","alpha","beta","intercept","sigma_int","sigma_slope","N","gamma","fit","fitnew")

#MCMC options
ni <- runs  # number of draws from the posterior
nt <- runs*.001   #thinning rate
nb <- runs*.2 # number to discard for burn-in
nc <- 2  # number of chains

#Jags

sim_niave = jags(inits=InitStage,
                 n.chains=nc,
                 model.file="Bayesian/NoDetect.jags",
                 working.directory=getwd(),
                 data=Dat,
                 parameters.to.save=ParsStage,
                 n.thin=nt,
                 n.iter=ni,
                 n.burnin=nb,
                 DIC=T)
```

```{r}
pars<-extract_par(sim_niave)
```

###Assess Convergence

```{r,cache=FALSE,eval=TRUE,fig.width=11,fig.height=5}

###Chains
ggplot(pars[pars$par %in% c("detect","alpha","beta"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + facet_grid(par~species,scale="free") + theme_bw() + labs(col="Chain") + ggtitle("Detection Probability")

```

```{r,fig.height=5,fig.width=11,eval=T}
ggplot(pars[pars$par %in% c("gamma","sigma_int","sigma_slope"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + theme_bw() + labs(col="Chain") + ggtitle("Trait-matching regression") + facet_wrap(~par,scales="free")
```

###Posteriors

```{r,cache=FALSE,fig.width=7,fig.height=5}
###Posterior Distributions
p<-ggplot(pars[pars$par %in% c("detect","alpha","beta"),],aes(x=estimate)) + geom_histogram() + ggtitle("Estimate of parameters") + facet_grid(species~par,scales="free") + theme_bw() + ggtitle("Species Posteriors")

#Add true values
tr<-melt(data.frame(species=1:length(detection),detect=detection,alpha=alpha,beta=beta),id.var='species')
colnames(tr)<-c("species","par","value")
psim<-p + geom_vline(data=tr,aes(xintercept=value),col='red',linetype='dashed',size=1)
psim
ggsave("Figures/SimulationPosteriorsNoDetect.jpg",dpi=300,height=8,width=8)
```

```{r,cache=FALSE,eval=TRUE,fig.height=3,fig.width=10}

p<-ggplot(pars[pars$par %in% c("gamma","polygamma","intercept","sigma_int","sigma_slope","sigma_polyslope"),],aes(x=estimate)) + geom_histogram() + ggtitle("Hierarchical Posteriors") + facet_grid(~par,scale="free") + theme_bw() 

#Add true values
tr<-melt(list(gamma=gamma,intercept=intercept,sigma_int=sigma_intercept,sigma_slope=sigma_slope))

colnames(tr)<-c("value","par")

psim2<-p + geom_vline(data=tr,aes(xintercept=value),linetype='dashed',size=1,col="red")
ggsave("Figures/SimulationHNoDetect.jpg",dpi=300,height=4,width=10)
```

**True values are given in the dashed lines.**

###Predicted Relationship 

```{r,fig.height=4,fig.width=4}

castdf<-group_by(pars,Chain) %>% sample_n(2000)%>% select(par,estimate) %>% filter(par %in% c("gamma","intercept"))

castdf<-dcast(pars[pars$par %in% c("gamma","intercept"),], Chain + Draw~par,value.var="estimate")

#calculated predicted y
predyniave<-trajF(alpha=castdf$intercept,beta=castdf$gamma,x=as.numeric(traitmatch))

orig<-trajF(alpha=rnorm(2000,intercept,sigma_intercept),beta=rnorm(2000,gamma,sigma_slope),x=as.numeric(traitmatch))

#save the pars to merge with other model
pars$Model<-"Without Detection"
pars_niave<-pars
```

##Hierarcichal Bayesian Model with detection

```{r,eval=T}
setwd("C:/Users/Ben/Documents/Occupy/")
runs<-20000

#Source model
source("Bayesian/NMixture.R")

#print model
print.noquote(readLines("Bayesian/NMixture.R"))

#Input Data
Dat <- list(
  Y=obs,
  Birds=dim(obs)[1],
  Plants=dim(obs)[2],
  Months=dim(obs)[3],
  traitmatch=traitmatch)

#A blank Y matrix - all present
initY<-array(dim=c(Dat$Birds,Dat$Plants),data=max(obs,na.rm=T))

#Inits
InitStage <- function() {list(beta=rep(.5,Dat$Birds),alpha=rep(.5,Dat$Birds),detect=rep(.5,Dat$Birds),intercept=0,tau_alpha=0.1,tau_beta=0.1,N=initY,gamma=0)}

#Parameters to track
ParsStage <- c("detect","alpha","beta","intercept","sigma_int","sigma_slope","N","gamma","fit","fitnew")

#MCMC options
ni <- runs  # number of draws from the posterior
nt <- runs*.001   #thinning rate
nb <- runs*.2 # number to discard for burn-in
nc <- 2  # number of chains

#Jags

m = jags(inits=InitStage,
         n.chains=nc,
         model.file="Bayesian/NMixture.jags",
         working.directory=getwd(),
         data=Dat,
         parameters.to.save=ParsStage,
         n.thin=nt,
         n.iter=ni,
         n.burnin=nb,
         DIC=T)
```

```{r}
pars<-extract_par(m)
```

###Assess Convergence

```{r,cache=FALSE,eval=TRUE,fig.width=11,fig.height=5}

###Chains
ggplot(pars[pars$par %in% c("detect","alpha","beta"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + facet_grid(par~species,scale="free") + theme_bw() + labs(col="Chain") + ggtitle("Detection Probability")

```

```{r,fig.height=5,fig.width=11,eval=T}
ggplot(pars[pars$par %in% c("gamma","sigma_int","sigma_slope"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + theme_bw() + labs(col="Chain") + ggtitle("Trait-matching regression") + facet_wrap(~par,scales="free")
```

###Posteriors

```{r,cache=FALSE,fig.width=7,fig.height=5}
###Posterior Distributions
p<-ggplot(pars[pars$par %in% c("detect","alpha","beta"),],aes(x=estimate)) + geom_histogram() + ggtitle("Estimate of parameters") + facet_grid(species~par,scales="free") + theme_bw() + ggtitle("Species Posteriors")

#Add true values
tr<-melt(data.frame(species=1:length(detection),detect=detection,alpha=alpha,beta=beta),id.var='species')
colnames(tr)<-c("species","par","value")
psim<-p + geom_vline(data=tr,aes(xintercept=value),col='red',linetype='dashed',size=1)
#ggsave("Figures/SimulationPosteriors.jpg",dpi=300,height=8,width=8)
```

```{r,cache=FALSE,eval=TRUE,fig.height=3,fig.width=10}

p<-ggplot(pars[pars$par %in% c("gamma","polygamma","intercept","sigma_int","sigma_slope","sigma_polyslope"),],aes(x=estimate)) + geom_histogram() + ggtitle("Hierarchical Posteriors") + facet_grid(~par,scale="free") + theme_bw() 

#Add true values
tr<-melt(list(gamma=gamma,intercept=intercept,sigma_int=sigma_intercept,sigma_slope=sigma_slope))

colnames(tr)<-c("value","par")

psim2<-p + geom_vline(data=tr,aes(xintercept=value),linetype='dashed',size=1,col="black")
#ggsave("Figures/SimulationH.jpg",dpi=300,height=4,width=10)
```

**True values are given in the dashed lines.**

###Compare estimates with and without detection

```{r,cache=FALSE,fig.width=7,fig.height=5}
pars$Model<-"Detection"

#Bind to other dataset
parsall<-rbind.data.frame(pars,pars_niave)
parsall$Model<-as.factor(parsall$Model)

###Posterior Distributions
p<-ggplot(parsall[parsall$par %in% c("detect","alpha","beta"),],aes(x=estimate,fill=Model)) + geom_histogram() + ggtitle("Estimate of parameters") + facet_grid(species~par,scales="free") + theme_bw() 

#Add true values
tr<-melt(data.frame(species=1:length(detection),detect=detection,alpha=alpha,beta=beta),id.var='species')
colnames(tr)<-c("species","par","value")
psim<-p + geom_vline(data=tr,aes(xintercept=value),col='black',linetype='dashed',size=1)
psim
#ggsave("Figures/SimulationPosteriorsBoth.jpg",dpi=300,height=8,width=8)
```

```{r,cache=FALSE,eval=TRUE,fig.height=3,fig.width=10}

p<-ggplot(parsall[parsall$par %in% c("gamma","polygamma","intercept","sigma_int","sigma_slope","sigma_polyslope"),],aes(x=estimate,fill=Model)) + geom_histogram() + ggtitle("Hierarchical Posteriors") + facet_grid(~par,scale="free") + theme_bw() 

#Add true values
tr<-melt(list(gamma=gamma,intercept=intercept,sigma_int=sigma_intercept,sigma_slope=sigma_slope))

colnames(tr)<-c("value","par")

psim2<-p + geom_vline(data=tr,aes(xintercept=value),linetype='dashed',size=1,col="black")
psim2
#ggsave("Figures/SimulationHBoth.jpg",dpi=300,height=4,width=10)
```

###Predicted Relationship 

Does accounting for non-independence and detection change our estimate of trait matching?

```{r,fig.height=4,fig.width=4}

castdf<-group_by(pars,Chain) %>% sample_n(2000)%>% select(par,estimate) %>% filter(par %in% c("gamma","intercept"))

castdf<-dcast(pars[pars$par %in% c("gamma","intercept"),], Chain + Draw~par,value.var="estimate")

trajF<-function(alpha,beta,x){
  indat<-cbind(alpha,beta)
  
  #fit regression for each input estimate
  sampletraj<-apply(indat,1,function(s){
    data.frame(x=x,y=exp(s['alpha'] + s['beta'] *x))})
  sample_all<-rbind_all(sampletraj)
  
  #Compute CI intervals
  predy<-group_by(sample_all,x) %>% summarise(lower=quantile(y,0.025,na.rm=T),upper=quantile(y,0.975,na.rm=T),mean=mean(y,na.rm=T))
  }

#calculated predicted y
predy<-trajF(alpha=castdf$intercept,beta=castdf$gamma,x=as.numeric(traitmatch))

orig<-trajF(alpha=rnorm(2000,intercept,sigma_intercept),beta=rnorm(2000,gamma,sigma_slope),x=as.numeric(traitmatch))


#plot and compare to original data
psim3<-ggplot(data=predy,aes(x=x)) + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.1,fill="red")  + geom_line(aes(y=mean),size=.8,col="red",linetype="dashed") + theme_bw() + ylab("Interactions") + geom_line(data=orig,aes(x=x,y=mean),col='black',size=1) + geom_ribbon(data=predyniave,aes(ymin=lower,ymax=upper),alpha=0.1,fill="blue") +  geom_line(data=predyniave,aes(y=mean),size=.8,col="blue",linetype="dashed") + xlab("Difference between Bill and Corolla Length")  
psim3
ggsave("Figures/SimulationResults.jpg",height=5,width=6,dpi=300)
```

Black line is the true relationship. The red line is the posterior mean with confidible intervals in shaded grey for the proposed bayesian model. The poisson glm that does not account for detectability or non-independence is in blue.

**Conclusion:** Accounting for detection and non-independence greatly increases the accuracy of the predicted state. The poisson glm underestimates the strength of trait matching among hummingbirds and their foodplants.

###Predicted Frequency

Validating the model

```{r,eval=T}

#any given state
predState<-group_by(parsall,par,Model) %>% filter(par=="N") %>% group_by(par,species,plant,Model) %>% summarize(State=as.numeric(names(which.max(table(estimate)))))

predState$species<-as.factor(predState$species)
predState$plant<-factor(predState$plant,levels=1:plant_species)

predplot<-ggplot(predState,aes(x=species,y=plant,fill=State)) + geom_tile() + scale_fill_continuous(na.value="white",low="white",high="red",limits=c(0,max(obs.state$Detection))) + labs(x="Hummingbird",y="Plant",fill="Predicted State") + ggtitle("Predicted") + facet_wrap(~Model)

#compare true and pred states
tr<-melt(true_interactions)
colnames(tr)<-c("Hummingbird","Plant","True_State")

mdat<-merge(predState,tr,by.x=c("species","plant"),by.y=c("Hummingbird","Plant"))

ggplot(mdat,aes(y=State,x=True_State,col=Model)) + geom_point(size=3) + geom_abline(linetype="dashed") + theme_bw() + xlab("True State") + ylab("Predicted State") + coord_equal()
```

###Posterior Check

Since I have simualted the data, it should fit as well as any random dataset drawn from the estimated parameters. An ideal fit would be posterior values sitting along the 1:1 line.

```{r,fig.height=4,fig.width=8}

fitstat<-parsall[parsall$par %in% c("fit","fitnew"),]
fitstat<-dcast(fitstat,Draw+Chain+Model~par,value.var="estimate")

#add 1:1 line

ymin<-round(min(c(fitstat$fit,fitstat$fitnew)))
ymax<-round(max(c(fitstat$fit,fitstat$fitnew)))
ab<-data.frame(x=ymin:ymax,y=ymin:ymax)
p<-ggplot(fitstat,aes(x=fit,y=fitnew)) + geom_point(aes(col=Model)) + theme_bw() + coord_equal()
psim4<-p  + labs(x="Discrepancy of observed data",y="Discrepancy of replicated data",col="Model") + geom_line(data=ab,aes(x=x,y=y)) + ggtitle("Simulated Data")

psim4

#Write Figures as panels
jpeg("Figures/SimulationPosteriorsBoth.jpg",height=12,width=12,units="in",res=300)
grid.arrange(psim,psim2,ncol=1,heights=c(3/4,1/4))
dev.off()

jpeg("Figures/SimPredictBoth.jpg",height=6,width=7,units="in",res=300)
grid.arrange(psim3)
dev.off()
```

#Observed dataset

```{r}
###Read in data
#read in flower morphology data, comes from Nectar.R
droppath<-"C:/Users/Ben/Dropbox/"
fl.morph<-read.csv(paste(droppath,"Thesis/Maquipucuna_SantaLucia/Results/FlowerMorphology.csv",sep=""))

#First row is empty
fl.morph<-fl.morph[-1,]

#Bring in Hummingbird Morphology Dataset, comes from
hum.morph<-read.csv(paste(droppath,"Thesis/Maquipucuna_SantaLucia/Results/HummingbirdMorphology.csv",sep=""))

#Bring in Interaction Matrix
int<-read.csv(paste(droppath,"Thesis/Maquipucuna_SantaLucia/Results/Network/HummingbirdInteractions.csv",sep=""),row.names=1)

#start with just cameras
#int<-int[int$Type=="Transect",]

#int$Month<-cut(int$Month,seq(0,12,2),labels=seq(1,11,2))

#Melt the interaction frame and match it with the traits
m.dat<-int[colnames(int) %in% c("ID","Video","Time","Hummingbird","Sex","TransectID","Transect_R","Iplant_Double","Pierce","DateP","Month","ele","Type")]

m.dat$Year<-years(as.Date(m.dat$DateP))
#Fix spacing to match clades

#Which are matching
levels(m.dat$Hummingbird)[!levels(m.dat$Hummingbird) %in% hum.morph$English]

#This needs to be checked
#print(paste(levels(m.dat$Hummingbird)[!levels(m.dat$Hummingbird) %in% hum.morph$English],"not matched"))

#Number of bird species
h_species<-nlevels(m.dat$Hummingbird)

#Number of plant species
plant_species<-nlevels(m.dat$Iplant_Double)

#levels(m.dat$Hummingbird)[!levels(m.dat$Hummingbird) %in% hum.morph$English]<-c("Green-crowned Woodnymph")
m.datH<-merge(m.dat,hum.morph, by.x="Hummingbird",by.y="English")

#Merge to flowers
int.FLlevels<-levels(factor(m.datH$Iplant_Double))

#Which flowers are we missing info for?
missingTraits<-int.FLlevels[!int.FLlevels %in% fl.morph$X]

#print(paste("Missing Trait Information:",missingTraits))
m.datH<-merge(m.datH,fl.morph, by.x="Iplant_Double",by.y="X")

#Drop piercing events, since they don't represent correlation
m.datH<-m.datH[!m.datH$Pierce %in% c("y","Y"),]

p<-ggplot(m.datH[!is.na(m.datH$Month),],aes(x=factor(Total_Culmen),TotalCorolla,col=Hummingbird)) + geom_point() 
p<-p + geom_smooth(aes(group=1),method="lm") + facet_wrap(~Month) + theme_bw()

p<-ggplot(m.datH[!is.na(m.datH$Month),],aes(x=Total_Culmen,TotalCorolla,col=Month)) + geom_point() 
p<-p + geom_smooth(aes(group=Month),method="lm") + theme_bw() + scale_color_continuous(low="blue",high="red")

#Difference Between Corolla and Bill Length of interactions measured
m.datH$BD<-abs(m.datH$Total_Culmen-m.datH$TotalCorolla)
p<-ggplot(m.datH,aes(y=BD,x=Hummingbird)) + geom_boxplot(position="dodge")
p<-p + coord_flip()

```

```{r}
###Format data

#remove species with less than  10 observations
keep<-names(which(table(m.datH$Hummingbird) > 10))

m.datH<-droplevels(m.datH[m.datH$Hummingbird %in% keep,])

#observed traitmatching
traitmatchF<-abs(t(sapply(hum.morph$Bill,function(x){x-fl.morph$TotalCorolla})))

rownames(traitmatchF)<-hum.morph$English
colnames(traitmatchF)<-fl.morph$Group.1
```

```{r}
#match names
#Round to 2 decimals
#Convert to cm for winbugs, avoids numerical underflow
traitmatchT<-round(traitmatchF[rownames(traitmatchF) %in% m.datH$Hummingbird,colnames(traitmatchF) %in% m.datH$Iplant_Double],2)/10
```

```{r}
#Formatting indexes
m.datH$Bird<-as.numeric(factor(m.datH$Hummingbird,sort(unique(m.datH$Hummingbird))))
m.datH$Plant<-as.numeric(factor(m.datH$Iplant_Double,sort(unique(m.datH$Iplant_Double))))
m.datH$obs<-1

traitmatchT<-traitmatchT[sort(rownames(traitmatchT)),sort(colnames(traitmatchT))]

#Create an index to connect to species names
#remerge species levels
bindex<-data.frame(Index=1:length(rownames(traitmatchT)),Species=rownames(traitmatchT))

pindex<-data.frame(Index=1:length(colnames(traitmatchT)),Species=colnames(traitmatchT))

indat<-filter(m.datH,!is.na(DateP)) %>% mutate(Time=DateP) %>% group_by(Bird,Plant,Time,Type) %>% summarize(Yobs=sum(obs)) 

indat$Time<-as.numeric(factor(indat$Time))

indatT<-dcast(indat,...~Type,value.var="Yobs")
#add in non-detection 0's
#this involves some ugly data management, where we have to figure out which plants were sampled in which periods, and if it was sampled, the non-detections are 0 if it wasn't the non-detections are NA. then remove all the Na's.

indatlong<-acast(indat[indat$Type=="Camera",],Bird~Plant~Time,value.var="Yobs",fill=0)
```


###Elevation ranges

Create a binary variable whether each observation was in a low elevation or high elevation transect. We have some species that just occur at the top of the gradient, and are not present in the sampling window of flowers at the low elevation.

```{r}

#1 is low elevation, #2 is high elevation, # 3 is both
#br<-c(1,3,3,2,2,3,2,3,1,2,2,1,3,3,3,1)
br<-c(3,2,2,2,3,2,1,3,3)
elevH<-data.frame(Species=keep,Range=br)
#merge to names
elevH<-merge(bindex,elevH)

#pr<-c(2,2,2,3,3,1,3,2,2,3,3,3,3,2,1,3,2,1,1,1,2,2,2,3,3,3,2,3,3,2,3,3,1,3,1)
pr<-c(2,3,3,3,3,2,3,2,3,2,3,2,1,1,2,2,2,3,3,3,2,3,3,2,3,3,1,3,1,2,1)

elevP<-data.frame(Species=colnames(traitmatchT),Range=pr)
elevP<-merge(pindex,elevP)

```

```{r}
#Only non-detections are real 0's, the rest are NA's and are removed.
#Plants not surveyed in that time period
#Hummingbirds not present at that site

for(x in 1:dim(indatlong)[3]){
  #Remove non sampled plants 
  a<-indatlong[,,x]
  toNA<-as.numeric(names(which(apply(a,2,sum)==0)))
  pres<-as.numeric(names(which(!apply(a,2,sum)==0)))
  
  indatlong[,toNA,x]<-NA
  
  #match elevation if there were presences, slightly terse code.
  if(length(pres)>0){
    for (y in 1:length(pres)){
      #Plant range
      pr<-elevP[elevP$Index %in% pres[y],"Range"]
      if(!pr==3){
        helim<-elevH[!elevH$Range %in% c(pr,3),"Index"]
        #print(paste(x,"=",sum(indatlong[helim,pres[y],x])))
        indatlong[helim,pres[y],x]<-NA
        }
      }  
    }
  }


#remove any 0's when a bird never feeds on a plant, the interaction must occur at some point to be observed.
for(x in 1:dim(indatlong)[1]){
  for(y in 1:dim(indatlong)[2]){
    a<-indatlong[x,y,]
    suma<-sum(a,na.rm=T)
    if(suma==0){
      indatlong[x,y,]<-NA
      }    
    }
  }

#melt and remove Na's
indat<-melt(indatlong)
indat<-indat[!is.na(indat$value),]

colnames(indat)<-c("Bird","Plant","Time","Yobs")
```

###Without detection

```{r,eval=T,results='hide'}

runs<-20000

#Source model
source("Bayesian/NoDetectRagged.R")

#print model
print.noquote(readLines("Bayesian/NoDetectRagged.R"))

#ragged index data

#Input Data
Dat <- list(
  Y=indat$Yobs,
  Bird=indat$Bird,
  Plant=indat$Plant,
  Birds=length(unique(indat$Bird)),
  Plants=max(indat$Plant),
  Nobs=length(indat$Yobs),
  traitmatch=traitmatchT
  )

#A blank Y matrix - all present
initY<-matrix(nrow=Dat$Birds,ncol=Dat$Plants,data=max(Dat$Y,na.rm=T))

#Inits
InitStage <- function() {list(beta=rep(.5,Dat$Birds),alpha=rep(.5,Dat$Birds),detect=rep(.999,Dat$Birds),gamma=0,intercept=0,tau_alpha=0.1,tau_beta=0.1,N=initY)}

#Parameters to track
ParsStage <- c("detect","alpha","beta","intercept","gamma","sigma_int","sigma_slope","N","fit","fitnew")

#MCMC options
ni <- runs  # number of draws from the posterior
nt <- runs*.001   #thinning rate
nb <- runs*.2 # number to discard for burn-in
nc <- 2  # number of chains

#Jags

m2_niave = jags(inits=InitStage,
                n.chains=nc,
                model.file="Bayesian/NoDetectRagged.jags",
                working.directory=getwd(),
                data=Dat,
                parameters.to.save=ParsStage,
                n.thin=nt,
                n.iter=ni,
                n.burnin=nb,
                DIC=T)
```


```{r}
pars_niave<-extract_par(m2_niave)
pars_niave$Model<-"Without Detection"
```


###Assess Convergence

```{r,cache=FALSE,eval=TRUE,fig.width=13,fig.height=5}

###Chains
ggplot(pars[pars$par %in% c("detect","alpha","beta"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + facet_grid(par~species,scale="free") + theme_bw() + labs(col="Chain") + ggtitle("Detection Probability")

```

```{r,fig.height=5,fig.width=11,eval=T}
ggplot(pars[pars$par %in% c("gamma","sigma_int","sigma_slope"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + theme_bw() + labs(col="Chain") + ggtitle("Trait-matching regression") + facet_wrap(~par,scales="free")
```

###With Detection

```{r,eval=T,results='hide'}

runs<-20000

#Source model
source("Bayesian/NMixtureRagged.R")

#print model
print.noquote(readLines("Bayesian/NMixtureRagged.R"))


#ragged index data

#Input Data
Dat <- list(
  Y=indat$Yobs,
  Bird=indat$Bird,
  Plant=indat$Plant,
  Birds=length(unique(indat$Bird)),
  Plants=max(indat$Plant),
  Nobs=length(indat$Yobs),
  traitmatch=traitmatchT
  )

#A blank Y matrix - all present
initY<-matrix(nrow=Dat$Birds,ncol=Dat$Plants,data=max(Dat$Y,na.rm=T))

#Inits
InitStage <- function() {list(beta=rep(.5,Dat$Birds),alpha=rep(.5,Dat$Birds),detect=rep(.5,Dat$Birds),gamma=0,intercept=0,tau_alpha=0.1,tau_beta=0.1,N=initY)}

#Parameters to track
ParsStage <- c("detect","alpha","beta","intercept","gamma","sigma_int","sigma_slope","N","fit","fitnew")

#MCMC options
ni <- runs  # number of draws from the posterior
nt <- runs*.001   #thinning rate
nb <- runs*.2 # number to discard for burn-in
nc <- 2  # number of chains

#Jags

m2 = jags(inits=InitStage,
          n.chains=nc,
          model.file="Bayesian/NMixtureRagged.jags",
          working.directory=getwd(),
          data=Dat,
          parameters.to.save=ParsStage,
          n.thin=nt,
          n.iter=ni,
          n.burnin=nb,
          DIC=T)
```

```{r}
pars<-extract_par(m2)

#name
pars$Model<-"Detection"
```

###Assess Convergence

```{r,cache=FALSE,eval=TRUE,fig.width=13,fig.height=5}

###Chains
ggplot(pars[pars$par %in% c("detect","alpha","beta"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + facet_grid(par~species,scale="free") + theme_bw() + labs(col="Chain") + ggtitle("Detection Probability")

```

```{r,fig.height=5,fig.width=11,eval=T}
ggplot(pars[pars$par %in% c("gamma","sigma_int","sigma_slope"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + theme_bw() + labs(col="Chain") + ggtitle("Trait-matching regression") + facet_wrap(~par,scales="free")
```

```{r}
#Bind together the two models
parsObs<-rbind(pars,pars_niave)
```

###Posteriors

```{r,cache=FALSE,fig.width=11,fig.height=12}
###Posterior Distributions
ggplot(parsObs[parsObs$par %in% c("detect","alpha","beta"),],aes(x=estimate,fill=Model)) + geom_histogram() + ggtitle("Estimate of parameters") + facet_grid(species~par,scales="free") + theme_bw() + ggtitle("Detection Probability")

```

```{r,cache=FALSE,fig.width=6,fig.height=7}
#Detection figure
ggplot(parsObs[parsObs$par %in% c("detect"),],aes(x=as.factor(species),y=estimate,fill=Model)) + geom_violin() + ggtitle("Estimate of parameters") + theme_bw() + ggtitle("Detection Probability") +facet_wrap(~Model,scales="free") 

parsp<-merge(pars,bindex,by.x="species",by.y="Index")
ggplot(parsp[parsp$par %in% c("detect"),],aes(x=estimate)) + geom_histogram() + ggtitle("Posterior Distribution") + theme_bw() + facet_wrap(~Species,ncol=4) + xlab("Probability of Detection")
ggsave("Figures/DetectionProb.jpg",dpi=300,height=7,width=11)
```


```{r,cache=FALSE,eval=TRUE,fig.height=5,fig.width=13}

ggplot(parsObs[parsObs$par %in% c("gamma","intercept","sigma_int","sigma_slope"),],aes(x=estimate,fill=Model)) + geom_histogram() + ggtitle("Trait matching regression parameters") + facet_wrap(~par,scale="free",nrow=2) + theme_bw() 
```

###Species Predictions

```{r}
#orignal data to plot as ppints
odat<-merge(indat,bindex,by.x="Bird",by.y="Index")
odat<-merge(odat,pindex,by.x="Plant",by.y="Index")

#Append the traits
odatT<-melt(traitmatchT)
colnames(odatT)<-c("Bird","Plant","x")
colnames(odat)<-c("Plant","Bird","Time","Yobs","Bird_Species","Plant_Species")
odat<-merge(odat,odatT,by.x=c("Bird_Species","Plant_Species"),by.y=c("Bird","Plant"))
#rename the column to match
colnames(odat)[1]<-"L1"
```

```{r,fig.height=10,fig.width=11,eval=F}

castdf<-dcast(parsObs[parsObs$par %in% c("beta","alpha","polybeta"),], species +Chain +Model+ Draw~par,value.var="estimate")

#Turn to 
castdf$species<-factor(castdf$species,levels=1:max(as.numeric(castdf$species)))

species.split<-split(castdf,list(castdf$species,castdf$Model))

species.traj<-list()

for(d in 1:length(species.split)){
  x<-species.split[[d]]
  index<-unique(x$species)
  species.traj[[d]]<-trajF(alpha=x$alpha,beta=x$beta,x=as.numeric(traitmatchT[index,]))
  }

names(species.traj)<-names(species.split)

species.traj<-melt(species.traj,id.var=colnames(species.traj[[1]]))

#split out names and model
species.traj[,c("Index","Model")]<-colsplit(species.traj$L1,"\\.",c("Index","Model"))

spe<-merge(species.traj,bindex,by.x="Index",by.y="Index")

#match colnames
names(odat)[1]<-"Species"
#plot and compare to original data
ggplot(data=spe,aes(x=x)) + geom_ribbon(aes(ymin=lower,ymax=upper,fill=Model),alpha=0.2)  + geom_line(aes(y=mean,col=Model),size=1) + theme_bw() + ylab("Interactions") + xlab("Difference between Bill and Corolla Length") + facet_wrap(~Species,scales="free",ncol=3) + labs(fill="Model") + geom_point(data=odat,aes(x=x,y=Yobs))

ggsave("Figures/SpeciesPredictionsBoth.jpg",dpi=300,height=8,width=10)
```

###Overall predicted relationship 

Does accounting for non-independence and detection change our estimate of trait matching?

```{r,fig.height=4,fig.width=4}

castdf<-dcast(parsObs[parsObs$par %in% c("gamma","intercept"),], Model+Chain + Draw~par,value.var="estimate")

castdf<-split(castdf,castdf$Model)
#calculated predicted y
predy<-rbind_all(lapply(castdf,function(i){
  #calculate trajectory and append model
  pr<-trajF(alpha=i$intercept,beta=i$gamma,x=round(as.numeric(traitmatchT),2))  
  pr$Model<-unique(i$Model)
  return(pr)
  }))

#plot and compare to original data
ggplot(data=predy,aes(x=x,col=Model,fill=Model)) + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.1)  + geom_line(aes(y=mean),size=1) + theme_bw() + ylab("Interactions") + xlab("Difference between Bill and Corolla Length") 
ggsave("Figures/ObsResults.jpg",dpi=300,height=5,width=7)

```

###Predicted Frequency

```{r,eval=T,fig.height=7,fig.width=12,eval=T}

#any given state
predState<-group_by(parsObs,par,Model) %>% filter(par=="N") %>% group_by(par,species,plant,Model) %>% summarize(State=as.numeric(names(which.max(table(estimate)))))

predState<-merge(predState,bindex,by.x="species",by.y="Index")

predState<-merge(predState,pindex,by.x="plant",by.y="Index")
colnames(predState)<-c("plant","species","par","Model","State","Bird","Plant")

#order it to show nestedness.
b_ord<-group_by(predState,Bird) %>% summarize(I=sum(State)) %>% arrange(desc(I)) %>% select(Bird)

p_ord<-group_by(predState,Plant) %>% summarize(I=sum(State)) %>% arrange(I) %>% select(Plant)

predState$Bird<-factor(predState$Bird,levels=b_ord$Bird)
predState$Plant<-factor(predState$Plant,levels=p_ord$Plant)

predplot<-ggplot(predState,aes(x=Bird,y=Plant,fill=State)) + geom_tile() + scale_fill_continuous(na.value="white",low="white",high="red",limits=c(0,max(obs.state$Detection))) + labs(x="Hummingbird",y="Plant",fill="Predicted State") + ggtitle("Predicted") + facet_wrap(~Model) + theme(axis.text.x = element_text(angle = 90))

orig<-group_by(indat,Bird,Plant) %>% summarize(Y=sum(Yobs))

finaldat<-merge(predState,orig,by.x=c("species","plant"),by.y=c("Bird","Plant"))
```

###Discrepancy 

The goodness if of fit is a measured as chi-squared. The expected value for each month is the detection rate * the estimate intensity of interactions. The expected value is compared to the observed value of the actual data. In addition, a replicate dataset is generated from the posterior predicted intensity. Better fitting models will have lower discrepancy values and be 
Better fitting models are smaller values and closer to the 1:1 line. A perfect model would be 0 discrepancy. This is unrealsitic given the stochasticity in the sampling processes. Rather, its better to focus on relative discrepancy. 

```{r,fig.height=4,fig.width=8}
fitstat<-parsObs[parsObs$par %in% c("fit","fitnew"),]
fitstat<-dcast(fitstat,Model+Draw+Chain~par,value.var="estimate")

ymin<-round(min(fitstat$fit))
ymax<-round(max(fitstat$fit))
ab<-data.frame(x=0:ymax,y=0:ymax)
disc_obs<-ggplot(fitstat,aes(x=fit,y=fitnew)) + geom_point(aes(col=Model)) + theme_bw() + labs(x="Discrepancy of observed data",y="Discrepancy of replicated data",col="Model")  + ggtitle("Field Observations") + geom_line(data=ab,aes(x=x,y=y)) + coord_fixed() + ylim(ymin=0,ymax=ymax) + xlim(xmin=0,xmax=ymax)
disc_obs
jpeg("Figures/ObsDiscrepancy.jpeg",res=300,height=7,width=6,units="in")
grid.arrange(psim4,disc_obs,nrow=2)
dev.off()
```

Tables

```{r}
#Detection table

dp<-group_by(pars[pars$par %in% c("detect"),],species) %>% summarise(mean=round(mean(estimate,na.rm=T),3)*100,lower=round(quantile(estimate,0.025,na.rm=T),3)*100,upper=round(quantile(estimate,0.975,na.rm=T),3)*100)

tab<-merge(dp,bindex,by.x="species",by.y="Index")[,-1]
tab[,c(4,1,2,3)]
write.csv(tab[,c(4,1,2,3)],"Figures/Table1.csv")
```

DIC Table
```{r}
DICtable<-data.frame(Data=c(rep("Simulated",2),rep("Observed",2)),Model=c(rep(c("No Detection Estimated","Detection Estimated"),2)),DIC=c(sim_niave$BUGSoutput$DIC,m$BUGSoutput$DIC,m2_niave$BUGSoutput$DIC,m2$BUGSoutput$DIC))
write.csv(dcast(DICtable,Data~Model),"Figures/DICtable.csv")
```

```{r}
save.image("Abundance.Rdata")
```
