---
title: "Hierarchical occupancy Models for species interactions: Empirical Data"
author: "Ben Weinstein - Stony Brook University"
output:
  html_document:
    toc: true
    number_sections: true
    theme: spacelab
  word_document: default
  keep_md: true
---

```{r,warning=FALSE,message=FALSE,echo=FALSE,cache=FALSE}
library(reshape2)
library(foreach)
library(doSNOW)
library(chron)
library(ggplot2)
library(knitr)
library(R2jags)
library(dplyr)
library(stringr)
library(gridExtra)
library(boot)
library(picante)
library(bipartite)

opts_chunk$set(message=FALSE,warning=FALSE,fig.width=5,fig.height=4,echo=TRUE,cache=F,fig.align='center',fig.path="figure/")

setwd("C:/Users/Ben/Documents/Occupy/")

set.seed(3)
#source functions

source("Functions.R")
```

```{r,echo=F,cache=FALSE}
paste("Run Completed at",Sys.time())
```

```{r}
#reload if needed
#load("Observed.Rdata")
```

#Observed dataset

```{r}
#read in flower morphology data, comes from Nectar.R
droppath<-"C:/Users/Ben/Dropbox/"
fl.morph<-read.csv(paste(droppath,"Thesis/Maquipucuna_SantaLucia/Results/FlowerMorphology.csv",sep=""))

#First row is empty
fl.morph<-fl.morph[-1,]

#Bring in Hummingbird Morphology Dataset, comes from
hum.morph<-read.csv(paste(droppath,"Thesis/Maquipucuna_SantaLucia/Results/HummingbirdMorphology.csv",sep=""))

#Bring in Interaction Matrix
int<-read.csv(paste(droppath,"Thesis/Maquipucuna_SantaLucia/Results/Network/HummingbirdInteractions.csv",sep=""),row.names=1)

#Correct known taxonomic disagreements, atleast compared to traits
int[int$Iplant_Double=="Alloplectus purpureus","Iplant_Double"]<-"Glossoloma purpureum"
int[int$Iplant_Double=="Capanea affinis","Iplant_Double"]<-"Kohleria affinis"
int[int$Iplant_Double=="Columnea cinerea","Iplant_Double"]<-"Columnea mastersonii"
int[int$Iplant_Double=="Alloplectus teuscheri","Iplant_Double"]<-"Drymonia teuscheri"
int[int$Iplant_Double=="Drymonia collegarum","Iplant_Double"]<-"Alloplectus tetragonoides"

#Some reasonable level of presences, 20 points
keep<-names(which(table(int$Hummingbird) > 25))

#sparkling violetear changes through time, seems to violate closed system assumption
keep<-keep[!keep %in% "Sparkling Violetear"]
int<-int[int$Hummingbird %in% keep, ]
dim(int)
m.dat<-droplevels(int[colnames(int) %in% c("ID","Video","Time","Hummingbird","Sex","TransectID","Transect_R","Iplant_Double","Pierce","DateP","Month","ele","Type")])

#Does the data come from camera or transect?
m.dat$Type<-(is.na(m.dat$TransectID))*1

m.dat$Year<-years(as.Date(m.dat$DateP))
#one missing date
m.dat$Year[m.dat$Year %in% 2012]<-2013

#Number of bird species
h_species<-nlevels(m.dat$Hummingbird)

#Number of plant species
plant_species<-nlevels(m.dat$Iplant_Double)

#Get english name
dath<-merge(m.dat,hum.morph, by.x="Hummingbird",by.y="English",keep=all)

#Merge to flowers
int.FLlevels<-levels(factor(dath$Iplant_Double))

#Which flowers are we missing info for?
missingTraits<-int.FLlevels[!int.FLlevels %in% fl.morph$X]

#print(paste("Missing Trait Information:",missingTraits))
dath<-merge(dath,fl.morph, by.x="Iplant_Double",by.y="X")

#Drop piercing events, since they don't represent correlation
dath<-dath[!dath$Pierce %in% c("y","Y"),]
```

##Match Species to Morphology

```{r}
#remove species with less than  10 observations
keep<-names(which(table(dath$Hummingbird) > 10))

dath<-droplevels(dath[dath$Hummingbird %in% keep,])

#observed traitmatching
traitmatchF<-abs(t(sapply(hum.morph$Bill,function(x){x-fl.morph$TotalCorolla})))

rownames(traitmatchF)<-hum.morph$English
colnames(traitmatchF)<-fl.morph$Group.1
```

```{r}
#match names #Round to 2 decimals #Convert to cm for winbugs, avoids numerical underflow
traitmatchT<-round(traitmatchF[rownames(traitmatchF) %in% dath$Hummingbird,colnames(traitmatchF) %in% dath$Iplant_Double],2)/10

traitmatchT<-traitmatchT[sort(rownames(traitmatchT)),sort(colnames(traitmatchT))]
```

##Elevation ranges

Create a binary variable whether each observation was in a low elevation or high elevation transect. We have some species that just occur at the top of the gradient, and are not present in the sampling window of flowers at the low elevation.

Accounting for non-availability.
We have to figure out which plants were sampled in which periods, and if it was sampled, the non-detection are 0 if it wasn't the non-detection are NA. then remove all the Na's.

```{r}
elevH<-read.csv("InputData/HummingbirdElevation.csv",row.names=1)
head(elevH)
colnames(elevH)[5]<-"Elevation"
elevH$Bird<-1:nrow(elevH)

#high elevation or low elevation
elevP<-read.csv("InputData/PlantElevation.csv",row.names=1)
colnames(elevP)[5]<-"Elevation"
elevP$Plant<-1:nrow(elevP)
elevP$Iplant_Double<-as.character(elevP$Iplant_Double)

#Correct known taxonomic errors
elevP[elevP$Iplant_Double %in% "Alloplectus purpureus","Iplant_Double"]<-"Glossoloma purpureum"
elevP[elevP$Iplant_Double %in% "Capanea affinis","Iplant_Double"]<-"Kohleria affinis"
elevP[elevP$Iplant_Double %in% "Alloplectus teuscheri","Iplant_Double"]<-"Drymonia teuscheri"
elevP[elevP$Iplant_Double %in% "Columnea cinerea","Iplant_Double"]<-"Columnea mastersonii"
elevP[elevP$Iplant_Double %in% "Alloplectus tenuis","Iplant_Double"]<-"Drymonia tenuis"

#Merge to observed Data
#plants
dathp<-merge(dath,elevP,by="Iplant_Double")

#birds
datph<-merge(dathp,elevH,by="Hummingbird")

```

### Summarize Observations

```{r}
#ID for NA is holger transects, make the id's 1:n for each day of transect at each elevation, assuming no elevation was split across days.
datph$ID<-as.character(datph$ID)

noid<-datph[is.na(datph$ID),]

id_topaste<-paste(noid$Transect_R,noid$DateP,"Transect",sep="_")
datph[which(is.na(datph$ID)),"ID"]<-id_topaste
  
indatraw<- datph %>% group_by(Bird,Plant,ID,DateP) %>% summarize(Yobs=n(),Elev=mean(ele,na.rm=T),Transect_R=unique(Transect_R)) 

indatraw[order(indatraw$Yobs,decreasing=T),]
#add unique ID
indatraw$Time<-1:nrow(indatraw)

#add months and years
indatraw$Month<-as.numeric(factor(months(strptime(indatraw$DateP,format="%Y-%m-%d")),levels=month.name))
indatraw$Year<-years(indatraw$DateP)

#Species names
for (x in 1:nrow(indatraw)){
  indatraw$Hummingbird[x]<-as.character(elevH[elevH$Bird %in% indatraw$Bird[x],"Hummingbird"])
  indatraw$Iplant_Double[x]<-as.character(elevP[elevP$Plant %in% indatraw$Plant[x],"Iplant_Double"])
}
```

What elevation transect is each observation in?
The camera data need to be inferred from the GPS point.

```{r}
#which elevation bin is each observation within
labs<-paste(seq(1300,2500,200),seq(1500,2700,200),sep="_")
indatraw[is.na(indatraw$Transect_R),"Transect_R"]<-cut(indatraw[is.na(indatraw$Transect_R),]$Elev,seq(1300,2700,200),labels=labs)
```

```{r}
#match the traits
traitmelt<-melt(traitmatchT)
colnames(traitmelt)<-c("Hummingbird","Iplant_Double","Traitmatch")
```

##Absences - accounting for non-detection

We have more information than just the presences, given species elevation ranges, we have absences as well. Absences are birds that occur at the elevation of the plant sample, but were not recorded feeding on the flower.

```{r}
indatlong<-acast(indatraw,Bird~Plant~Time,value.var="Yobs")
indatlong[is.na(indatlong)]<-0
```

```{r}
#Only non-detections are real 0's, the rest are NA's and are removed.
#Plants not surveyed in that time period
#Hummingbirds not present at that site

for(x in 1:dim(indatlong)[3]){
  #Remove non sampled plants 
  a<-indatlong[,,x]
  
  #No observations at that plant
  toNA<-as.numeric(names(which(apply(a,2,sum)==0)))
  pres<-as.numeric(names(which(!apply(a,2,sum)==0)))
  indatlong[,colnames(a) %in% toNA,x]<-NA
  
  #match elevation if there were presences, slightly terse code.
  if(length(pres)>0){
    for (y in 1:length(pres)){
      #Plant range
      pr<-elevP[elevP$Plant %in% pres[y],"Elevation"]
      #if there is not enough elevation area, make NA
      if(is.na(pr)|!pr==3){
        helim<-elevH[!elevH$Elevation %in% c(pr,3),"Bird"]
        #sanity check, make sure rows don't have values
        sums<-indatlong[rownames(a) %in% helim,colnames(a) %in% pres[y],x]==0
        helim<-helim[sums]
        #set to NA if they are outside of elevation and are blank
        indatlong[rownames(a) %in% helim ,colnames(a) %in% pres[y],x]<-NA
        }
      }  
    }
  }

#melt and remove Na's
indat<-melt(indatlong)
indat<-indat[!is.na(indat$value),]

colnames(indat)<-c("Bird","Plant","Time","Yobs")
```

```{r}
#remerge the time period data
Timelookup<-indatraw %>% dplyr::select(Time,DateP,Transect_R,Month,Year)%>% group_by(Time,DateP,Transect_R,Month,Year) %>% distinct() %>% arrange(Time)

#Get time information
indat<-merge(indat,Timelookup[,!colnames(Timelookup) %in% c("Bird","Plant")],by=c("Time"),all.x=T)

#Species names
for (x in 1:nrow(indat)){
  indat$Hummingbird[x]<-as.character(elevH[elevH$Bird %in% indat$Bird[x],"Hummingbird"])
  indat$Iplant_Double[x]<-as.character(elevP[elevP$Plant %in% indat$Plant[x],"Iplant_Double"])
}

#Get trait information
#match the traits
indat<-merge(indat,traitmelt,by=c("Hummingbird","Iplant_Double"))
```

Reformat index for jags.
Jags needs a vector of input species 1:n with no breaks.

```{r}
indat$Hummingbird<-as.factor(indat$Hummingbird)
indat$Iplant_Double<-as.factor(indat$Iplant_Double)
indat$jBird<-as.numeric(indat$Hummingbird)
indat$jPlant<-as.numeric(indat$Iplant_Double)

jagsIndexBird<-data.frame(Hummingbird=levels(indat$Hummingbird),jBird=1:length(levels(indat$Hummingbird)))
 
jagsIndexPlants<-data.frame(Iplant_Double=levels(indat$Iplant_Double),jPlant=1:length(levels(indat$Iplant_Double)))
```

Similiarly, the trait matrix needs to reflect this indexing.

# Hierarcichal Occupancy Model

$$ Y_{i,j,k} \sim Binom(N_{i,j},detect_i)$$
$$N_{i,j} \sim Pois(\lambda_{i,j}) $$
$$log(\lambda_{i,j})<-\alpha_i + \beta_i * abs(Bill_i - Corolla_i) $$
$$detect_i \sim N(dprior,\tau_detect)$$     

**Priors**

$$\alpha_i \sim N(intercept,\tau_{\alpha})$$
$$\beta_i \sim N(\gamma,\tau_{detect})$$
$$ logit(detect_i) \sim N(dprior,tau_detect)$$

**Hyperpriors**
$$dprior \sim N(0,0.1)$$
$$gamma \sim N(0,0.0001)$$
$$intercept \sim N(0,0.0001)$$

$$\tau_{\alpha} \sim Gamma(0.0001,0.0001)$$
$$\tau_\beta \sim Gamma(0.0001,0.0001)$$
$$\tau_detect \sim Gamma(0.0001,0.0001)$$

**Derived quantities**

$$\sigma_{intercept} = \sqrt[2]{\frac{1}{\tau_\alpha}}$$
$$\sigma_{slope} = \sqrt[2]{\frac{1}{\tau_\beta}}$$
$$\sigma_{detect} = \sqrt[2]{\frac{1}{\tau_detect}}$$

# Poisson GLM

```{r,eval=T}
runs<-60000

#trigger parallel
paralleljags<-T

#Source model
source("Bayesian/NoDetectNmixturePoissonRagged.R")

#print model
print.noquote(readLines("Bayesian//NoDetectNmixturePoissonRagged.R"))

if(paralleljags){

  #for parallel run
  Yobs=indat$Yobs
  Bird=indat$jBird
  Plant=indat$jPlant
  Time=indat$Time
  Traitmatch=indat$Traitmatch
  #number of birds to iterate
  Birds=max(indat$jBird)
  Nobs=length(indat$Yobs)

  #A blank Y matrix - all present
  Ninit<-rep(max(indat$Yobs)+1,Nobs)
  
  #Inits
  InitStage <- function() {list(beta=rep(0.5,Birds),alpha=rep(0.5,Birds),intercept=0,tau_alpha=0.1,tau_beta=0.1,N=Ninit,gamma=0)}
  
  #Parameters to track
  ParsStage <- c("alpha","beta","intercept","sigma_int","sigma_slope","ynew","gamma","fit","fitnew")
  
  #MCMC options
  ni <- runs  # number of draws from the posterior
  nt <- 1   #thinning rate
  nb <- runs*.95 # number to discard for burn-in
  nc <- 2  # number of chains

  Dat<-list("Yobs","Bird","Plant","Time","Traitmatch","Birds","Nobs","Ninit")

    m2_niave<-do.call(jags.parallel,list(Dat,InitStage,ParsStage,model.file="Bayesian/NoDetectNmixturePoissonRagged.jags",n.thin=nt, n.iter=ni,n.burnin=nb,n.chains=nc))
    
  } else {
 #Input Data
  Dat <- list(
    Yobs=indat$Yobs,
    Bird=indat$jBird,
    Plant=indat$jPlant,
    Time=indat$Time,
    Traitmatch=indat$Traitmatch,
    #number of birds to iterate
    Birds=max(indat$jBird),
    Nobs=length(indat$Yobs))
  
    #A blank Y matrix - all present
    Ninit<-rep(max(Dat$Yobs)+1,Dat$Nobs)
  
    #Inits
  InitStage <- function() {list(beta=rep(0.5,Dat$Birds),alpha=rep(0.5,Dat$Birds),dtrans=rep(0,Dat$Birds),intercept=0,tau_alpha=0.1,tau_beta=0.1,N=Ninit,gamma=0)}
  
  #Parameters to track
  ParsStage <- c("detect","alpha","beta","intercept","sigma_int","sigma_slope","ynew","gamma","fit","fitnew","dprior","sigma_detect")
  
  #MCMC options
  ni <- runs  # number of draws from the posterior
  nt <- 2   #thinning rate
  nb <- runs*.95 # number to discard for burn-in
  nc <- 2  # number of chains
  
  #Jags
  
  m2_niave <- jags(inits=InitStage,
                   n.chains=nc, model.file="Bayesian/NoDetectNmixturePoissonRagged.jags",
                   working.directory=getwd(),
                   data=Dat,
                   parameters.to.save=ParsStage,
                   n.thin=nt,
                   n.iter=ni,
                   n.burnin=nb,
                   DIC=T)
}
```


```{r,eval=F}
#recompile if needed
load.module("dic")
runs<-50000
recompile(m2_niave)
m2_niave<-update(m2_niave,n.iter=runs,n.burnin=runs*.75)
```

```{r}
pars_dniave<-extract_par(m2_niave,data=indat,Bird="jBird",Plant="jPlant")
pars_dniave$Model<-"Poisson GLM"
```

##Assess Convergence

```{r,cache=FALSE,fig.width=13,fig.height=5}
###Chains
ggplot(pars_dniave[pars_dniave$par %in% c("detect","alpha","beta"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + facet_grid(par~species,scale="free") + theme_bw() + labs(col="Chain") + ggtitle("Detection Probability")
```

```{r,fig.height=5,fig.width=11}
ggplot(pars_dniave[pars_dniave$par %in% c("gamma","sigma_int","sigma_slope"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + theme_bw() + labs(col="Chain") + ggtitle("Trait-matching regression") + facet_wrap(~par,scales="free")
```

# Observed Data With Detection

```{r,eval=T}
runs<-500000

#trigger parallel
paralleljags<-T

#Source model
source("Bayesian/NmixturePoissonRagged.R")

#print model
print.noquote(readLines("Bayesian//NmixturePoissonRagged.R"))

if(paralleljags){

  #for parallel run
  Yobs=indat$Yobs
  Bird=indat$jBird
  Plant=indat$jPlant
  Time=indat$Time
  Traitmatch=indat$Traitmatch
  #number of birds to iterate
  Birds=max(indat$jBird)
  Nobs=length(indat$Yobs)

  #A blank Y matrix - all present
  Ninit<-rep(max(indat$Yobs)+1,Nobs)
  
  #Inits
  InitStage <- function() {list(beta=rep(0.5,Birds),alpha=rep(0.5,Birds),dtrans=rep(0,Birds),intercept=0,tau_alpha=0.1,tau_beta=0.1,N=Ninit,gamma=0)}
  
  #Parameters to track
  ParsStage <- c("detect","alpha","beta","intercept","sigma_int","sigma_slope","ynew","gamma","fit","fitnew","dprior","sigma_detect")
  
  #MCMC options
  ni <- runs  # number of draws from the posterior
  nt <- 5   #thinning rate
  nb <- runs*.95 # number to discard for burn-in
  nc <- 2  # number of chains

  Dat<-list("Yobs","Bird","Plant","Time","Traitmatch","Birds","Nobs","Ninit")

    system.time(m2<-do.call(jags.parallel,list(Dat,InitStage,ParsStage,model.file="Bayesian/NmixturePoissonRagged.jags",n.thin=nt, n.iter=ni,n.burnin=nb,n.chains=nc)))
  } else {
  #Input Data
  Dat <- list(
    Yobs=indat$Yobs,
    Bird=indat$jBird,
    Plant=indat$jPlant,
    Time=indat$Time,
    Traitmatch=indat$Traitmatch,
    #number of birds to iterate
    Birds=max(indat$jBird),
    Nobs=length(indat$Yobs))
  
    #A blank Y matrix - all present
    Ninit<-rep(max(Dat$Yobs)+1,Dat$Nobs)
  
    #Inits
  InitStage <- function() {list(beta=rep(0.5,Dat$Birds),alpha=rep(0.5,Dat$Birds),dtrans=rep(0,Dat$Birds),intercept=0,tau_alpha=0.1,tau_beta=0.1,N=Ninit,gamma=0)}
  
  #Parameters to track
  ParsStage <- c("detect","alpha","beta","intercept","sigma_int","sigma_slope","ynew","gamma","fit","fitnew","dprior","sigma_detect")
  
  #MCMC options
  ni <- runs  # number of draws from the posterior
  nt <- 10   #thinning rate
  nb <- runs*.95 # number to discard for burn-in
  nc <- 2  # number of chains
  
  #Jags
  
  system.time(m2 <- jags(inits=InitStage,
                   n.chains=nc, model.file="Bayesian/NmixturePoissonRagged.jags",
                   working.directory=getwd(),
                   data=Dat,
                   parameters.to.save=ParsStage,
                   n.thin=nt,
                   n.iter=ni,
                   n.burnin=nb,
                   DIC=T))
}
```


```{r,eval=F}
#recompile if needed
load.module("dic")
runs<-1000000
recompile(m2)
m2<-update(m2,n.iter=runs,n.burnin=runs*.95,n.thin=10)
```

```{r}
#extract par to data.frame
pars_detect<-extract_par(m2,data=indat,Bird="jBird",Plant="jPlant")

#name
pars_detect$Model<-"Detection"
```

##Assess Convergence

```{r,cache=FALSE,fig.width=13,fig.height=5}
###Chains
ggplot(pars_detect[pars_detect$par %in% c("detect","alpha","beta"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + facet_grid(par~species,scale="free") + theme_bw() + labs(col="Chain") + ggtitle("Detection Probability")
```

###Hierarcichal Posteriors

```{r,fig.height=5,fig.width=11}
ggplot(pars_detect[pars_detect$par %in% c("gamma","intercept","sigma_int","sigma_slope","dprior","sigma_detect"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + theme_bw() + labs(col="Chain") + ggtitle("Trait-matching regression") + facet_wrap(~par,scales="free")
ggsave("Figures/BothObs.svg",height=5,width=7)
ggsave("Figures/BothObs.jpg",height=5,width=7,dpi=300)
```

```{r}
#Bind together the two models
parsObs<-rbind(pars_detect,pars_dniave)
```

##Posteriors

```{r,cache=FALSE,fig.width=11,fig.height=14}
###Posterior Distributions
ggplot(parsObs[parsObs$par %in% c("detect","alpha","beta"),],aes(x=estimate,fill=Model)) + geom_histogram() + ggtitle("Estimate of parameters") + facet_grid(species~par,scales="free") + theme_bw() + ggtitle("Detection Probability")
```

```{r,cache=FALSE,fig.width=10,fig.height=5}
#Detection figure
ggplot(parsObs[parsObs$par %in% c("detect"),],aes(x=as.factor(species),y=estimate,fill=Model)) + geom_violin() + ggtitle("Estimate of parameters") + theme_bw() + ggtitle("Detection Probability") +facet_wrap(~Model,scales="free") 

pars_detect<-merge(pars_detect,jagsIndexBird,by.x="species",by.y="jBird")
ggplot(pars_detect[pars_detect$par %in% c("detect"),],aes(x=estimate)) + geom_histogram() + ggtitle("Posterior Distribution") + theme_bw() + facet_wrap(~Hummingbird,ncol=5) + xlab("Probability of Detection")
ggsave("Figures/DetectionProb.jpg",dpi=300,height=7,width=11)
```

```{r,cache=FALSE,fig.height=5,fig.width=13}
ggplot(parsObs[parsObs$par %in% c("gamma","intercept","sigma_int","sigma_slope","dprior"),],aes(x=estimate,fill=Model)) + geom_histogram() + ggtitle("Trait matching regression parameters") + facet_wrap(~par,scale="free",nrow=2) + theme_bw() 
```

##Species Predictions

```{r,fig.height=10,fig.width=11}
castdf<-dcast(parsObs[parsObs$par %in% c("beta","alpha"),], species +Chain +Model+ Draw~par,value.var="estimate")

#Turn to species level
castdf$species<-factor(castdf$species,levels=1:max(as.numeric(castdf$species)))

species.split<-split(castdf,list(castdf$species,castdf$Model))

species.traj<-list()

for(d in 1:length(species.split)){
  x<-species.split[[d]]
  #species name
  index<-jagsIndexBird[unique(x$species),"Hummingbird"]
  
  #range of trait distances
  tsp<-indat %>% filter(Hummingbird==index) %>% distinct(Traitmatch) %>% .$Traitmatch
  species.traj[[d]]<-trajF(alpha=x$alpha,beta=x$beta,x=tsp)
  }

names(species.traj)<-names(species.split)

species.traj<-melt(species.traj,id.var=colnames(species.traj[[1]]))

#split out names and model
species.traj[,c("Index","Model")]<-colsplit(species.traj$L1,"\\.",c("Index","Model"))

spe<-merge(species.traj,jagsIndexBird,by.x="Index",by.y="jBird")

#match colnames

#plot and compare to original data
ggplot(data=spe[,],aes(x=x)) + geom_point(data=indat,aes(x=Traitmatch,y=Yobs)) + geom_ribbon(aes(ymin=lower,ymax=upper,fill=Model),alpha=0.2)  + geom_line(aes(y=mean,col=Model),size=1) + theme_bw() + ylab("Interactions") + xlab("Difference between Bill and Corolla Length") + facet_wrap(~Hummingbird,scales="free",ncol=3)+ labs(fill="Model") 

ggsave("Figures/SpeciesPredictionsBoth.jpg",dpi=300,height=8,width=10)
```

###Overall predicted relationship 

Does accounting for non-independence and detection change our estimate of trait matching?

```{r,fig.height=6,fig.width=10}

castdf<-dcast(parsObs[parsObs$par %in% c("gamma","intercept"),], Model+Chain + Draw~par,value.var="estimate")

castdf<-split(castdf,castdf$Model)
#calculated predicted y
predy<-rbind_all(lapply(castdf,function(i){
  #calculate trajectory and append model
  pr<-trajF(alpha=i$intercept,beta=i$gamma,x=indat$Traitmatch)  
  pr$Model<-unique(i$Model)
  return(pr)
  }))

#plot and compare to original data
ggplot(data=predy,aes(x=x,col=Model,fill=Model)) + geom_point(data=indat,aes(x=Traitmatch,y=Yobs),col="black",fill="black",size=2) + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.6)  + geom_line(aes(y=mean),size=1) + theme_bw() + ylab("Interactions") + xlab("Difference between Bill and Corolla Length") 
ggsave("Figures/ObsResults.jpg",dpi=300,height=5,width=7)
```

##Predicted Frequency

```{r,fig.height=7,fig.width=12}
#any given state
predState<-group_by(parsObs,par,Model) %>% filter(par=="ynew") %>% group_by(par,species,plant,Model) %>% summarize(State=as.numeric(names(which.max(table(estimate)))))

#order it to show nestedness.
b_ord<-group_by(predState,species) %>% summarize(I=sum(State)) %>% arrange(desc(I)) %>% select(species)

p_ord<-group_by(predState,plant) %>% summarize(I=sum(State)) %>% arrange(I) %>% select(plant)

predState$Bird<-factor(predState$species,levels=b_ord$species)
predState$Plant<-factor(predState$plant,levels=p_ord$plant)

predplot<-ggplot(predState,aes(x=Bird,y=Plant,fill=State)) + geom_tile() + scale_fill_continuous(na.value="white",low="white",high="red",limits=c(0,max(indat$Yobs))) + labs(x="Hummingbird",y="Plant",fill="Predicted State") + ggtitle("Predicted") + facet_wrap(~Model) + theme(axis.text.x = element_text(angle = 90))
predplot
orig<-group_by(indat,Bird,Plant) %>% summarize(Y=sum(Yobs))

finaldat<-merge(predState,orig,by.x=c("species","plant"),by.y=c("Bird","Plant"))
```

##Discrepancy 

The goodness if of fit is a measured as chi-squared. The expected value for each day is the detection rate * the estimate intensity of interactions. The expected value is compared to the observed value of the actual data. In addition, a replicate dataset is generated from the posterior predicted intensity. Better fitting models will have lower discrepancy values and be 
Better fitting models are smaller values and closer to the 1:1 line. A perfect model would be 0 discrepancy. This is unrealsitic given the stochasticity in the sampling processes. Rather, its better to focus on relative discrepancy. In addition, a model with 0 discrepancy would likely be seriously overfit and have little to no predictive power.

```{r,fig.height=4,fig.width=8,message=F,warning=F}
fitstat<-parsObs[parsObs$par %in% c("fit","fitnew"),]
fitstat<-dcast(fitstat,Model+Draw+Chain~par,value.var="estimate")

ymin<-round(min(fitstat$fit))
ymax<-round(max(fitstat$fit))
ab<-data.frame(x=0:ymax,y=0:ymax)
disc_obs<-ggplot(fitstat,aes(x=fit,y=fitnew)) + geom_point(aes(col=Model)) + theme_bw() + labs(x="Discrepancy of observed data",y="Discrepancy of replicated data",col="Model")  + ggtitle("Empirical Data") + geom_line(data=ab,aes(x=x,y=y)) + coord_fixed() + ylim(ymin=0,ymax=max(max(c(fitstat$fit,fitstat$fitnew)))) + xlim(xmin=0,xmax=max(max(c(fitstat$fit,fitstat$fitnew)))) + facet_wrap(~Model,scales = "free")
disc_obs
```

##Detection table

```{r}
dp<-group_by(pars_detect[pars_detect$par %in% c("detect"),],species) %>% summarise(mean=round(mean(estimate,na.rm=T),3)*100,lower=round(quantile(estimate,0.025,na.rm=T),3)*100,upper=round(quantile(estimate,0.975,na.rm=T),3)*100)

tab<-merge(dp,jagsIndexBird,by.x="species",by.y="jBird")[,-1]
tab[,c(4,1,2,3)]
write.csv(tab[,c(4,1,2,3)],"Figures/Table1.csv")
```

Number of samples per species

##DIC Table

```{r}
DICtable<-data.frame(Data=c(rep("Simulated",2),rep("Observed",2)),Model=c(rep(c("No Detection Estimated","Detection Estimated"),2)),DIC=c(sim_niave$BUGSoutput$DIC,sim_detect$BUGSoutput$DIC,m2_niave$BUGSoutput$DIC,m2$BUGSoutput$DIC))
h<-dcast(DICtable,Data~Model)
print(h)
write.csv(h,"Figures/DICtable.csv")
```

#Predicted versus Observed Data

```{r}
m<-max(traitmatchT)-traitmatchT

mat<-indat %>% group_by(jBird,jPlant) %>% summarize(n=sum(Yobs))
true_state<-acast(mat,jBird~jPlant,fill=0)

#get species names?

#trait liklihood
dmultinom(true_state,prob=m,log=T)
paste("Correlation coefficient is:", round(cor(c(true_state),c(m),method="spearman"),2))
```

###Test Statistic
Chisquared statistic is (observed-expected)^2/(expected + 0.5), we add the 0.5 to avoid dividing by 0. For each combination of birds and plants, predicted versus  observed for each data point.

##Multinomial prediction

Based on trait differences.

```{r}
#define discrep function
chisq<-function(o,e){(o-e)^2/(e+0.5)}

nullm<-function(){
r<-mgen(m,sum(true_state),keep.species = F)
}

#same number of draws as bayesian
cl<-makeCluster(20,"SOCK")
registerDoSNOW(cl)
mult_matrix<-foreach(x=1:length(pars_detect[pars_detect$par %in% "fit","estimate"]),.packages=c("bipartite","reshape2")) %dopar% nullm()
stopCluster(cl)

mats<-lapply(mult_matrix,function(r){
rmerge<-matrix(nrow = nrow(true_state),ncol=ncol(true_state))
colnames(rmerge)<-colnames(true_state)
  rownames(rmerge)<-rownames(true_state)
#for each position what is the chisq
for (x in 1:nrow(r)){
  for (y in 1:ncol(r)){
   rmerge[x,y]<-chisq(o=r[x,y],e=true_state[x,y])
  }
}

#sum of chisq driscrepancy
return(rmerge)
})

multi_disc<-sapply(mats,function(x) sum(x))

qplot(multi_disc)+ xlab("Chisquared Discrepancy for Multimonial Liklihood") + geom_vline(xintercept=mean(multi_disc),col='red',linetype='dashed')
```

#Compare Bayesian Occupancy and Multinomial using true known interactions

## Poisson GLM

```{r}
N_niave<-pars_dniave[ pars_dniave$par %in% "ynew",]

bydraw<-split(N_niave,list(N_niave$Chain,N_niave$Draw))

occ_nodetect_matrix<-lapply(bydraw,function(x){
  r<-acast(x,species ~ plant,value.var = "estimate",fun.aggregate = sum)
})

#calculate discrep on those deviates
occ_nodetect<-lapply(occ_nodetect_matrix,function(r){
    #for each position what is the chisq
  rmerge<-matrix(nrow = nrow(true_state),ncol=ncol(true_state))
  colnames(rmerge)<-colnames(true_state)
  rownames(rmerge)<-rownames(true_state)
  for (x in 1:nrow(r)){
    for (y in 1:ncol(r)){
     rmerge[x,y]<-chisq(o=r[x,y],e=true_state[x,y])
      }
    }
  return(rmerge)
})

names(occ_nodetect)<-1:length(occ_nodetect)
names(occ_nodetect_matrix)<-1:length(occ_nodetect_matrix)
```

##With Detection

```{r}
N<-pars_detect[pars_detect$par %in% "ynew",]
bydraw<-split(N,list(N$Chain,N$Draw))
occ_matrix<-lapply(bydraw,function(x){
  r<-acast(x,species ~ plant,value.var = "estimate",fun.aggregate = sum)
  })

#calculate discrep for those aggregated matrices
occ<-lapply(occ_matrix,function(r){
    #for each position what is the chisq
  rmerge<-matrix(nrow = nrow(true_state),ncol=ncol(true_state))
  colnames(rmerge)<-colnames(true_state)
  rownames(rmerge)<-rownames(true_state)
  for (x in 1:nrow(r)){
    for (y in 1:ncol(r)){
     rmerge[x,y]<-chisq(o=r[x,y],e=true_state[x,y])
      }
  }
  return(rmerge)
})

names(occ)<-1:length(occ)
names(occ_matrix)<-1:length(occ_matrix)

```

##Compare to observed data

```{r}
mmat<-melt(true_state)
colnames(mmat)<-c("Bird","Plant","True_State")

#append to predicted matrices

#occupancy with detection
mocc<-melt(occ_matrix)
colnames(mocc)<-c("Bird","Plant","Occupancy","Iteration")
simdat<-merge(mocc,mmat,by=c("Bird","Plant"))

#occupancy with nodetection
moccd<-melt(occ_nodetect_matrix)
colnames(moccd)<-c("Bird","Plant","Poisson GLM","Iteration")

simdat<-merge(simdat,moccd,by=c("Bird","Plant","Iteration"))

#multinomial
multmats<-melt(mult_matrix)
colnames(multmats)<-c("Bird","Plant","Multinomial","Iteration")
simdat<-merge(simdat,multmats,by=c("Bird","Plant","Iteration"))


simdat<-melt(simdat,measure.vars = c("Detect","Poisson GLM","Multinomial"))

ggplot(simdat,aes(x=True_State,y=value,col=variable)) + geom_point() + geom_abline() + labs(col="Model") + ylab("Predicted State") + xlab("True State") + theme_bw() + facet_wrap(~variable)


ggplot(data=simdat[simdat$Bird %in% 13 & simdat$Plant %in% 40,],aes(x)) + geom_histogram(aes(x=value,fill=variable)) + geom_vline(aes(xintercept=unique(True_State))) + facet_wrap(~Plant,scales='free')
```

##Summary of discrepancy of predicted matrices

```{r}
#Multinomial
multi_disc<-sapply(mats,function(x) mean(x))

#occupancy without detection
occno_disc<-sapply(occ_nodetect,function(x) mean(x))

#Occupancy with detection
occ_disc<-sapply(occ,function(x) mean(x))

#compared to bayesian
ggplot(data.frame(multi_disc)) + geom_histogram(aes(x=multi_disc),fill="blue",alpha=.6)+ xlab("Chi-squared Discrepancy") + geom_histogram(data=data.frame(occ_disc),aes(x=occ_disc),fill="red",alpha=.6) + theme_bw() +geom_vline(aes(xintercept=mean(occ_disc)),linetype="dashed",col="red")+ geom_vline(xintercept=mean(multi_disc),linetype="dashed",col="blue") + geom_histogram(data=data.frame(occno_disc),aes(x=occno_disc),fill="orange",alpha=.6) + geom_vline(aes(xintercept=mean(occno_disc)),linetype="dashed",col="orange")
```

##Comparison of summary statistics for all three approaches

```{r}
d<-list(Occupancy=occ,Multinomial=mats,Poisson_GLM=occ_nodetect)
d<-melt(d)
colnames(d)<-c("Bird","Plant","value","Iteration","Model")

d %>% group_by(Model,Iteration) %>% summarize(mean=mean(value),sd=sd(value),sum=sum(value)) %>% group_by(Model) %>% summarize(mean_mean=round(mean(mean),2),mean_sd=round(sd(mean),2),mean_sum=round(mean(sum),2))
```

```{r}
save.image("Observed.Rdata")
```
