---
title: "AbundanceMixture"
author: "Ben Weinstein"
date: "Monday, March 16, 2015"
output: html_document
---

**In this study we want to estimate three quantitites:**

$$\hat{Y}_{i,j} = \text{Abundance of Interactions between Hummingbird i  at  Flower j}$$

$$ \psi_{i,j} = \text{the probability a flower is visited by a hummigbird}  $$

$$ p_{i,j} = \text{the probability of detecting a bird-flower interaction given that it occurs} $$


#Simulation: Constant detection for each hummingbird speices, Interaction frequency is a function of species hierarcichal model and corolla similarity. Uneven sampling among species.

### Parameters

  * 7 hummingbird species
    * Range of bill sizes (mm) Pois(10)
  * Ten plants
    * Range of corolla sizes (mm) Pois(10)
  * Mean occupancy ($\gamma$) for each hummingbird is drawn from U(0.5,1)  
    * For each plant the occupancy is N($\gamma$,0.2) (truncated,0-1) 
    * Strong trait matching (minimizing Bill-Corolla difference)
      $logit(true_state)<-\alpha + \beta *traitmatch$
      where $\alpha=0, \beta = 1$
      
      * Keep it linear for now (only penalty for having increasingly smaller bill) 
  * Imperfect detection $p = U(0,1)$ for each hummingbird species
  * 24 month replicates
  * Phenology = .75 (plants are in flower 3/4 of surveys)

```{r,fig.height=5,fig.width=8}
#Number of hummingbird species
h_species=4
plant_species=20
months=24
detection=runif(h_species,0,1)
phenology=0.5

#Bill sizes
Bill<-rpois(h_species,10)

#Corolla sizes
Corolla<-rpois(plant_species,20)

#Subtract both
traitmatch<-sapply(Corolla,function(x) x - Bill)

#regression slope
intercept=0.5
gamma=-0.2
sigma_slope=0.02
sigma_intercept=0.02

beta<-rnorm(h_species,gamma,sigma_slope)
alpha<-rnorm(h_species,intercept,sigma_intercept)

#fit regression
occupancy<-inv.logit(alpha + beta * traitmatch)

df<-dcast(melt(list(occ=occupancy,traitmatch=traitmatch)),Var1+Var2~L1)
ggplot(df,aes(x=traitmatch,y=occ,col=as.factor(Var1))) + geom_point(size=6) + geom_smooth(method="lm",aes(group=Var1)) + theme_bw() + labs(x="Bill-Corolla Difference",y="Occurrence Probability","Hummingbird") + ggtitle("Correlation in Simulated Data")

#create data matrix
dat<-matrix(nrow=h_species,ncol=plant_species,data=rbinom(occupancy,1,occupancy))

```

### True Interaction Matrix

```{r}
#Reshape into a nicer format
mdat<-melt(dat)
colnames(mdat)<-c("Plant","Hummingbird","True_State")
trueplot<-ggplot(mdat,aes(x=as.factor(Plant),y=as.factor(Hummingbird),fill=as.factor(True_State))) + geom_tile() + scale_fill_manual(labels=c(0,1),values=c("White","Black")) + labs(x="Hummingbird",y="Plant",fill="Presence") + ggtitle("True State ")
```

```{r,fig.width=7,fig.height=3}
##Simulate detection 

#for each species loop through and create a replicate dataframe
obs<-array(dim=c(h_species,plant_species,months))

#Which months are plants in bloom?
#This is the phenology matrix
sampled<-replicate(plant_species,rbinom(months,1,phenology))
sampled[sampled==0]<-NA

for(x in 1:h_species){
  for (y in 1:plant_species){
    ts<-dat[x,y]
    
    #Detections
    d<-rbinom(months,1,detection[x])
    
    #detections are function of phenology detection probability and occupancy
    obs[x,y,]<-d *sampled[,y] * ts
    }
  }

#plot observed state
obs.state<-melt(obs)
colnames(obs.state)<-c("Hummingbird","Plant","Month","Presence")
obs.state<-group_by(obs.state,Hummingbird,Plant) %>% summarize(Detections=sum(Presence,na.rm=T))

obplot<-ggplot(obs.state,aes(x=Hummingbird,y=as.factor(Plant),fill=Detections)) + geom_tile() + scale_fill_continuous(low="white",high="Red") + ylab("Plant") + ggtitle("Observed")
grid.arrange(trueplot,obplot,nrow=1)
```

## Occupancy Model Formulation

$$ Y_{i,j,k} \sim Bern(sightp_{i,j,k})$$
$$sightp_{i,j,k} = present_{i,j} * detect_i$$
$$present_{i,j} \sim Bern(occ_{i,j}) $$
$$occ_{i,j}<-\alpha + \beta * (Bill_i - Corolla_i) $$

```{r,eval=T}
setwd("C:/Users/Ben/Documents/Occupy/Bayesian")

runs<-30000

#Source model
source("Simulation5.R")
 
#print model
print.noquote(readLines("Simulation5.R"))

#Input Data
Dat <- list(
  Y=obs,
  Plants=plant_species,
  Birds=h_species,
  Months=months,
  traitmatch=traitmatch
  )

#A blank Y matrix - all present
initY<-matrix(nrow=h_species,ncol=plant_species,data=1)

#Inits
InitStage <- function() {list(beta=rep(.5,h_species),alpha=rep(.5,h_species),detect=rep(.5,h_species),gamma=0,intercept=0,tau_alpha=0.1,tau_beta=0.1,present=initY)}

#Parameters to track
ParsStage <- c("occ","detect","alpha","beta","intercept","gamma","sigma_int","sigma_slope","present")

#MCMC options
ni <- runs  # number of draws from the posterior
nt <- runs*.001   #thinning rate
nb <- 0 # number to discard for burn-in
nc <- 2  # number of chains

#Jags

m5 = jags(inits=InitStage,
         n.chains=nc,
         model.file="Simulation5.jags",
         working.directory=getwd(),
         data=Dat,
         parameters.to.save=ParsStage,
         n.thin=nt,
         n.iter=ni,
         n.burnin=nb,
         DIC=T)
```

```{r,echo=FALSE,eval=F}
#Update if needed
recompile(m5)
add<-1000000
m5 <- update(m5,n.iter=add)
```

```{r}
  #extract desired info from the models
  parsO<-melt(m5$BUGSoutput$sims.array)
  colnames(parsO)<-c("Draw","Chain","parameter","estimate")
  parsO<-parsO[!parsO$Draw %in% 1:(max(parsO$Draw)-2000/m5$BUGSoutput$n.chains),]
  
  #label species and plants
  l<-levels(parsO$parameter)

  #parameters to save
  totrack<-m5$parameters.to.save
  
  sp_pl<-data.frame(parameter=l,str_match(l,pattern="(\\d+),(\\d+)")[,-1],par=str_extract(l,paste(totrack,collapse="|")))
  
  colnames(sp_pl)<-c("parameter","species","plant","par")

  sp_pl[is.na(sp_pl$species),c("species")]<-str_extract(sp_pl[is.na(sp_pl$species),"parameter"],"\\d+")
  pars<-merge(parsO,sp_pl)
  
  #take out deviance
  pars<-pars[!pars$par %in% "deviance",]
  
```

###Assess Convergence

```{r,cache=FALSE,eval=TRUE,fig.width=8,fig.height=4}

###Chains
ggplot(pars[pars$par %in% c("detect","alpha","beta"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + facet_grid(par~species,scale="free") + theme_bw() + labs(col="Chain") + ggtitle("Detection Probability")

```

```{r,fig.height=2,fig.width=12,eval=F}
ggplot(pars[pars$par %in% c("alpha","beta","sigma_int","sigma_slope"),],aes(x=Draw,y=estimate,col=as.factor(Chain))) + geom_line() + theme_bw() + labs(col="Chain") + ggtitle("Occupancy Probability") + facet_wrap(~par,scales="free")
```

###Posteriors

```{r,cache=FALSE,fig.width=9,fig.height=3}
###Posterior Distributions
p<-ggplot(pars[pars$par %in% c("detect"),],aes(x=estimate)) + geom_histogram() + ggtitle("Estimate of parameters") + facet_wrap(~species) + theme_bw() + ggtitle("Detection Probability")

#Add true values
tr<-melt(data.frame(species=1:length(detection),detect=detection),id.var='species')
colnames(tr)<-c("species","par","value")
p + geom_vline(data=tr,aes(xintercept=value),col='red',linetype='dashed',size=1)
```

```{r,cache=FALSE,eval=TRUE,fig.height=3,fig.width=10}

p<-ggplot(pars[pars$par %in% c("gamma","intercept","sigma_int","sigma_slope"),],aes(x=estimate)) + geom_histogram() + ggtitle("Trait matching regression parameters") + facet_grid(~par,scale="free") + theme_bw() 

#Add true values
tr<-melt(list(gamma=gamma,intercept=intercept,sigma_int=sigma_intercept,sigma_slope=sigma_slope))

colnames(tr)<-c("value","par")

p + geom_vline(data=tr,aes(xintercept=value),linetype='dashed',size=1,col="red")
```

**True values are given in the dashed lines.**

###Predicted Relationship
```{r,fig.height=4,fig.width=4}
 
castdf<-group_by(pars,Chain) %>% sample_n(1000)%>% select(par,estimate) %>% filter(par %in% c("gamma","intercept"))

castdf<-dcast(pars[pars$par %in% c("gamma","intercept"),], Chain + Draw~par,value.var="estimate")
castdf<-castdf[sample(1:nrow(castdf),1000),]

trajF<-function(alpha,beta,x){
  indat<-cbind(alpha,beta)
  
  #fit regression for each input estimate
  sampletraj<-apply(indat,1,function(s){
    data.frame(x=x,y=inv.logit(s['alpha'] + s['beta'] *x))})
  sample_all<-rbind_all(sampletraj)
  
  #Compute CI intervals
  predy<-group_by(sample_all,x) %>% summarise(lower=quantile(y,0.025),upper=quantile(y,0.975),mean=mean(y))
  }

#calculated predicted y
predy<-trajF(alpha=castdf$intercept,beta=castdf$gamma,x=as.numeric(traitmatch))

orig<-trajF(alpha=rnorm(2000,intercept,sigma_intercept),beta=rnorm(2000,gamma,sigma_slope),x=as.numeric(traitmatch))

#plot and compare to original data
ggplot(data=predy,aes(x=x)) + geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.2)  + geom_line(aes(y=mean),size=1.1,col="red") + theme_bw() + ylab("Probability of Occurrence") + geom_ribbon(data=orig,aes(x=x,ymin=lower,ymax=upper),fill='blue',alpha=.5) + xlab("Difference between Bill and Corolla Length")
```

Blue area is the true relationship with known variance. The red line is the mean estimated relationship with confidence intervals in shaded grey. 

###Predicted Occupancy

```{r}

#any given state

predState<-group_by(pars,par) %>% filter(par=="present") %>% group_by(par,species,plant) %>% summarize(State=as.numeric(names(which.max(table(estimate)))))

predState$species<-as.factor(predState$species)
predState$plant<-factor(predState$plant,levels=1:plant_species)

predplot<-ggplot(predState,aes(x=species,y=plant,fill=as.factor(State))) + geom_tile() + scale_fill_manual(labels=c("Absent","Present"),values=c("white","black")) + labs(x="Hummingbird",y="Plant",fill="Predicted State") + ggtitle("Predicted")

grid.arrange(trueplot,obplot,predplot,nrow=1)
```
